Paper Discussion Report

### 2022-04-07 paper_discussion_report/paper_discussion.tex
Fixed some typos in paper discussion report

  %
  % Here we have the typical use of a "T" for an initial drop letter
  % and "HIS" in caps to complete the first word.
  \IEEEPARstart{A}{} paper authored by Gunasekaran, et al. uses deep learning methods 
  to classify viruses by their DNA sequences[1]. Since DNA is composed of strings
  of nucleotides, the problem amounts to classifying viruses according to samples of nucleotide
  strings. The authors collect data from the public nucleotide sequence database,
  The National Centre for Biotechnology Information (NCBI) https://www.ncbi.nlm.nih.gov.
  They then encode this data using label encoding and kmer encoding. For each encoding type,
  they run three different deep learning models: CNN, CNN-LSTM, and CNN-Bidirectional-LSTM.
  The architectures of all three models start with embedding layers, then convolutional layers,
  then max pooling layers, and then from there diverge to either add LSTM layers or 
  bidirectional LSTM layers before finishing with dense layers and a final output layer.
  The authors compare all six combinations of the two encoding methods with the three model types
  using several performance metrics. They also compare their results to current state-of-the-art
  approaches in the literature.

@@

\section{Data Preprocessing}
  The authors encoded the data in two different formats for comparative analysis.
  In the first approach, they use label encoding, which replaces each nucleoid by a unique index value,
  preserving positional information (Fig 4).
  In the second approach, they used kmer encoding, which generates all kmers from a sequence 
  and forms an English-like sentence onto which natural language processing techniques
  can be applied (Fig 5).

@@

  \begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/kmer_encoding.png}
    \caption{Kmer encoding example, where k=6, borrowed from Gunasekaran, H., et al. (2020)}
  \end{figure}

@@

\section{Proposed Models}
  In this paper, the authors use three different classification models to perform multi-classification 
  of the DNA sequence: CNN, CNN-LSTM and CNN-bidirectional LSTM.
  All three models share the same initial layers: an embedding layer, convolutional layers, and
  max pooling layers. The raw data is first either label encoded or kmer encoded and then 
  one-hot encoded to be fed into the embedding layer. The results of the embedding layer
  are then fed into the convolutional layers, each of which has a corresponding max pooling layer.
  From here, the three model types diverge. For the CNN model, the results are fed into
  the final dense layers and output layers. The LSTM and bidirectional LSTM models differ only
  in that they add LSTM or bidirectional LSTM respectively before the final dense layers.

@@

  The classification models are evaluated using different classification metrics like accuracy, precision, recall, F1 score, sensitivity, 
  and specificity by obtaining the confusion matrix for both k-mer and label encoding techniques.
  The confusion matrix holds the values of true positives (TP), true negatives (TN), false positives (FP), and false negatives(FN). 
  Based on these values, classification metrics are calculated and compared.
  Table III shows the formulas for calculating these classification metrics. 
  \begin{table}
    \caption{\label{tab:table3}Formulas to Calculate Performance Metrics}
    \centering
    % \resizebox{\columnwidth}{!}{%
      \begin{tabular}{ || l l ||}
      \hline
      Metric & Formula \\ 
      \hline
      \hline
      Accuracy    & $\frac{TP + TN}{TP + TN + FP + FN}$\\  
      Specificity & $\frac{TN}{TN + FP}$ \\
      Sensitivity & $\frac{TP}{TP + FN}$ \\
      Precision   & $\frac{TP}{TP + FP}$ \\
      \hline
      \end{tabular}%
    % }
  \end{table}

>>>>

  %
  % Here we have the typical use of a "T" for an initial drop letter
  % and "HIS" in caps to complete the first word.
  \IEEEPARstart{A}{} paper authored by Gunasekaran, et al. uses deep learning methods
  to classify viruses by their DNA sequences[1]. Since DNA is composed of strings
  of nucleotides, the problem amounts to classifying viruses according to samples of nucleotide
  strings. The authors collect data from the public nucleotide sequence database,
  The National Centre for Biotechnology Information (NCBI) https://www.ncbi.nlm.nih.gov.
  They then encode this data using label encoding and k-mer encoding. For each encoding type,
  they run three different deep learning models: CNN, CNN-LSTM, and CNN-Bidirectional-LSTM.
  The architectures of all three models start with embedding layers, then convolutional layers,
  then max pooling layers, and then from there diverge to either add LSTM layers or
  bidirectional LSTM layers before finishing with dense layers and a final output layer.
  The authors compare all six combinations of the two encoding methods with the three model types
  using several performance metrics. They also compare their results to current state-of-the-art
  approaches in the literature.

@@

\section{Data Preprocessing}
  The authors encoded the data in two different formats for comparative analysis.
  In the first approach, they use label encoding, which replaces each nucleoid by a unique index value,
  preserving positional information (Fig 4).
  In the second approach, they used k-mer encoding, which generates all k-mers from a sequence
  and forms an English-like sentence onto which natural language processing techniques
  can be applied (Fig 5).

@@

  \begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/kmer_encoding.png}
    \caption{K-mer encoding example, where k=6, borrowed from Gunasekaran, H., et al. (2020)}
  \end{figure}

@@

\section{Proposed Models}
  In this paper, the authors use three different classification models to perform multi-classification
  of the DNA sequence: CNN, CNN-LSTM and CNN-bidirectional LSTM.
  All three models share the same initial layers: an embedding layer, convolutional layers, and
  max pooling layers. The raw data is first either label encoded or k-mer encoded and then
  one-hot encoded to be fed into the embedding layer. The results of the embedding layer
  are then fed into the convolutional layers, each of which has a corresponding max pooling layer.
  From here, the three model types diverge. For the CNN model, the results are fed into
  the final dense layers and output layers. The LSTM and bidirectional LSTM models differ only
  in that they add LSTM or bidirectional LSTM respectively before the final dense layers.

@@

  The classification models are evaluated using different classification metrics like accuracy, precision, recall, F1 score, sensitivity,
  and specificity by obtaining the confusion matrix for both k-mer and label encoding techniques.
  The confusion matrix holds the values of true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN).
  Based on these values, classification metrics are calculated and compared.
  Table III shows the formulas for calculating these classification metrics.
  \begin{table}
    \caption{\label{tab:table3}Formulas to Calculate Performance Metrics}
    \centering
    % \resizebox{\columnwidth}{!}{%
      \begin{tabular}{ || l l ||}
      \hline
      Metric & Formula \\
      \hline
      \hline
      Accuracy    & $\frac{TP + TN}{TP + TN + FP + FN}$\\
      Specificity & $\frac{TN}{TN + FP}$ \\
      Sensitivity & $\frac{TP}{TP + FN}$ \\
      Precision   & $\frac{TP}{TP + FP}$ \\
      \hline
      \end{tabular}%
    % }
  \end{table}
